Conservation of interaction partners between related plants varies widely across communities and between plant families.

### 2019-04-18 manuscript/phylogenetic_signal_manuscript.tex
all major comments addressed, I think. Now need to read and update

\author{Alyssa R. Cirtwill$^{1}$, Giulio V. Dalla Riva$^{2}$, Nick J. Baker$^{1}$,\\
 Joshua A. Thia$^{1,3}$, Christie J. Webber$^{1}$, Daniel B. Stouffer$^{1}$, MSC helpers from Linkoping}
\date{\small$^1$Centre for Integrative Ecology, School of Biological Sciences\\
    \medskip$^2$Biomathematics Research Centre, School of Mathematics and Statistics\\
            University of Canterbury\\Private Bag 4800\\
Christchurch 8140, New Zealand\\
\medskip$^3$Present Address: School of Biological Sciences\\
University of Queensland\\Brisbane, QLD 4072, Australia }

@@

  [[update refs, add definition of plant roles]]
  If attractive and/or defensive traits are heritable,
  then we can reasonably expect that related plants will have similar 
  patterns of interactions with animals~\citep{Schemske1999}.
  Recent studies that have investigated this question at the level of whole
  communities, however, have yielded mixed results~\citep{Rezende2007a,Gomez2010,Rohr2014a,Fontaine2015,Lind2015}.
  In particular, significant phylogenetic signal in plants' interaction partners
  ---the tendency for more closely-related plants to have more similar interactions---
  tends to be rare in empirical networks (\citealp{Rezende2007a,Lind2015}; but see~\citealp{Elias2013,Fontaine2015}).
  Further, plants' roles [[how defined?]] within networks tend to be less phylogenetically constrained than those of 
  animals~\citep{Rezende2007a,Chamberlain2014,Rohr2014,Vamosi2014,Lind2015}. [[GVDR is skeptical of this - check partiteness, size, type of role and add more detail to support.]]

@@

  We then estimated phylogenetic distances between species. To accomplish 
  this, we constructed a phylogenetic tree for our dataset based on a dated
  `mega-tree' of angiosperms~\citep{Zanne2014}. Some species were not included
  in the angiosperm mega-tree (largely ferns and tree ferns). For angiosperms,
  a sister taxon was identified using~\citet{APW} and the species added manually.
  Ferns, tree ferns, and a single club moss were added to the base of the tree.
  This means that closely-related non-angiosperm species appear to have very long 
  phylogenetic distances between them. For this reason, we excluded comparisons 
  between pairs of non-angiosperms from our analyses. As only two networks (both 
  plant-herbivore networks) included more than one such species and non-angiosperms
  were always a small minority of any network, we do not 
  believe that omitting these comparisons has greatly affected our results.
  To obtain trees for each network, we 
  pruned the dated mega-tree to include only species in that network.

@@

  % [[It appears we switched to Jaccard and Sorenson dissimilarities - number of shared partners and number of shared partners*2/notshared.]]
\subsection*{Statistical analysis}

  We modelled the relationship between niche overlap and phylogenetic distance
  using a logistic regression. We used both the numbers of shared ($M_{ij}$)and 
  non-shared ($U_{ij}$) partners as dependent variables and centred, scaled phylogenetic 
  distance as the independent variable. This approach is conceptually similar to 
  modelling successes and failures in a binomial-distributed process. 
  Accordingly, we assumed a binomial distribution for residuals. Regressions of
  niche overlap and phylogenetic distance within each network were fit using the
  R~\citep{R} base function ``glm'' and took the form. The fixed effect of distance
  in this regression, $\beta_{distance}$, can
  be understood as the change in log odds of sharing an interaction partner per
  million-year change in phylogenetic distance.

  These separate regressions avoid the potential
  for confounding the effects of different relationships in different networks. As
  we also wished to evaluate the overall trend across networks, we fit an additional
  regression of niche overlap and phylogenetic distance across all network types.
  As well as the fixed effect of phylogenetic distance, this regression included 
  fixed effects of network type (pollination or herbivory) and the interaction 
  between phylogenetic network type and random intercepts and slopes per network.
  This expanded regression was fit using the R~\citep{R} function ``glmer'' from
  package \emph{lme4}~\citep{lme4}.

@@

@@

  For each plant family, within each network type, we then fit one of two similar 
  sets of models. If family $f$ was found in several networks of the same type
  (e.g., several pollination networks), we fit a mixed-effects logistic regression
  relating niche overlap to a fixed effect of phylogenetic distance and a random
  effect for each network using the R~\citep{R} function ``glmer'' from
  package \emph{lme4}~\citep{lme4}. If family $f$ was found in only one network
  (and therefore only one network type), we omitted the 
  network-level random effect and fit a logistic regression using the R~\citep{R}
  base function ``glm''. [[We didn't use equations to explain the first round
  of logistic regressions, so not sure why I was doing that here...]]

@@

  \subsection*{Within-network conservation of niche overlap} [[ok, need to make some updates here.]]

    Across all networks, overlap of interaction partners decreased with increasing phylogenetic distance and the decrease in overlap of pollinators was steeper than that in herbivores ($\beta_{distance:herbivory}$=-6.82, $p$\textless0.001 and $\beta_{distance:pollination}$=-18.5, $p$\textless0.001, respectively). That is, a pair of plants in the same genus was more likely to share interaction partners than a pair of plants in the same family in both types of networks, but a pair of congeners would be less likely to share pollinators than to share herbivores. Regardless of phylogenetic distance, a pair of plants were also less likely to share pollinators than herbivores ($\beta_{PP}$=-1.44, $p$\textless0.001). This may be due to the greater proportion of
    specialist pollinators than specialist herbivores. In our dataset, an
    average of 48\% (+/- 14) of pollinators in a given web were extreme 
    specialists (i.e., visited only one plant species) compared to 29\% 
    (+/- 29) of herbivores ($z$=5.62, df=68, $P$\textless0.001 
    for a binomial regression of specialists and generalists over network
    type). % Are the above betas un-scaled?

    Despite these general trends, there was substantial variation between 
    pollination networks, with overlap of interaction partners decreasing 
    with increasing phylogenetic distance in some networks and increasing in 
    others (Fig.~\ref{within_network_regression}). Overlap of interaction partners decreased significantly with increasing phylogenetic distance in 7/11 plant-herbivore networks and 33/59 plant-pollinator networks. In the remaining four plant-herbivore networks and 25 of the 26 remaining plant-pollinator networks, overlap of interaction partners was not related to phylogenetic distance. Overlap of interaction partners increased with increasing phylogenetic distance in only a single plant-pollinator network.% (M_PL_026).

    [[Reviewers demanded this - it seems very boring]]
    Comparing the results in the observed networks to those obtained after permuting phylogenetic distances across pairs of plants, the observed slope of the relationship between phylogenetic distance and interaction partner overlap was always more extreme (i.e., always lesser or always greater) than that obtained in the permuted networks (Fig.~\ref{obs_vs_random}). Observed networks with a negative relationship between phylogenetic distance and overlap always had a more negative slope than that obtained from the permuted networks, while the 10 networks with positive relationships between phylogenetic distance and overlap always had more positive relationships than the permuted networks. This suggests that even in the networks with non-significant relationships between niche overlap and phylogenetic distance, the association between niche overlap and phylogenetic distance was not random.

    Comparing the permuted networks to permutations of the permuted networks, the slope obtained from the initial permuted network was quite variable with respect to the slopes obtained from 500 permutationss of the permuted network. Averaged over the 1000 permutations of each observed network, the slope of the permuted network was more extreme than 48.1-51.3\% of the permutations of the permuted network. This confirms that shuffling phylogenetic distances between plant pairs destroys the relationship between distance and interaction partner overlap, and that further shuffling distances does not have a predictable effect.

@@

  % Regression of all families at once with RE's for network
    Taking all families together, the probability of species in the same family sharing interaction partners decreased wiht increasing phylogenetic distance in both plant-pollinator and plant-herbivore networks and the strength of this relationship did not differ significantly between network types ($\beta_{distance}$=-4.75, $p$=0.012; $\beta_{distance:PH}$=-1.73, $p$=0.681). Plants in plant-herbivore networks had a higher probability of sharing interaction partners with a member of the same family ($\beta_{PH}$=0.78, $p$=0.007).

    The relationship between within-family niche overlap and phylogenetic distance
    varied widely in both pollination and herbivory networks. 
    % Gotta separate plant-herb and plant-poll networks.

    % Regressions for each family separately.
    For the families
    that were well represented in pollination networks, overlap decreased
    with increasing phylogenetic distance in 18 (Table~\ref{family_slopes_pp}).
    There was no significant relationship between overlap and phylogenetic distance in
    a further 15 plant families. Finally, the overlap between pairs of \emph{Polygonaceae}
    increased with increasing phylogenetic distance. 
    Of the seven plant families that were sufficiently well represented in herbivory 
    networks, overlap decreased with increasing phylogenetic distance in four 
    (Table~\ref{family_slopes_ph}; Fig.~\ref{within_family_regression}). Two
    families did not show significant relationships between phylogenetic distance and overlap,
    and in one family, \emph{Fabaceae}, overlap of interaction partners increased with 
    increasing phylogenetic distance.

    Models for two families did not converge. In both the \emph{Lauraceae}, (represented by four species in one pollination network) and the \emph{Sapindaceae} in plant-herbivore networks (represented by five species in one network), one pair of species shared a single interaction partner while all other pairs did not share any interaction partners. The model for \emph{Sapindaceae} in plant-pollinator networks did converge.

@@

  ARC, DBS, GVDR, and NJB designed the research. ARC, JAT, and CJW collected published data. [[Add MSC helpers!]]
  ARC and GVDR performed the analyses. All authors contributed to the manuscript.

@@

  \begin{figure}[!h]
    \begin{center}
      \centerline{\includegraphics*[width=.75\textwidth]{Figures/dataplots/observed_vs_random.eps}}
    \end{center}
     \caption{The slopes of the mixed-effect logistic regression of pairwise niche overlap against phylogenetic distance was signficantly different from 0 for each network. Here we show, for plant-herbivore (PH) and plant-pollinator (PP) networks, the observed slopes (depicted by green squares or purple diamonds, respectively). The maximum and minimum slopes obtained from 999 permutations of each network are depicted by thin, black lines. For both network types, the slopes obtained from permuted networks were always very close to 0.}
    \label{obs_vs_random} 
  \end{figure}

>>>>

\author{Alyssa R. Cirtwill$^{1,2}$, Giulio V. Dalla Riva$^{3}$, Nick J. Baker$^{1}$,\\
Mikael Ohlsson$^{4}$, Isabelle Norstr\"{o}m$^{4}$, Inger-Marie Wohlfarth Hasle$^{4}$,\\
Joshua A. Thia$^{1,5}$, Christie J. Webber$^{1}$, Daniel B. Stouffer$^{1}$}
\date{\small$^1$Centre for Integrative Ecology, School of Biological Sciences\\
\medskip$^2$Department of Ecology,\\
Environment, and Plant Sciences (DEEP)\\
Stockholm University\\
114 19 Stockholm, Sweden\\
\medskip$^3$Biomathematics Research Centre, School of Mathematics and Statistics\\
University of Canterbury\\Private Bag 4800\\
Christchurch 8140, New Zealand\\
\medskip$^4$Department of Physics, Chemistry, and Biology (IFM)\\ Link\"{o}ping University\\ 581 83 Link\"{o}ping, Sweden
\medskip$^5$Present Address: School of Biological Sciences\\
University of Queensland\\Brisbane, QLD 4072, Australia }

@@

  [[update refs, perhaps expand?]]
  If attractive and/or defensive traits are heritable,
  then we can reasonably expect that related plants will have similar 
  patterns of interactions with animals~\citep{Schemske1999}. That is, there may be \emph{phylogenetic signal} in plants' interactions such that closely-related plants may tend to have similar interaction partners.
  Recent studies that have investigated this question at the level of whole
  communities, however, have yielded mixed results~\citep{Rezende2007a,Gomez2010,Rohr2014a,Fontaine2015,Lind2015}.
  In particular, significant phylogenetic signal in plants' sets of interaction partners tends to be rare in empirical networks (\citealp{Rezende2007a,Lind2015}; but see~\citealp{Elias2013,Fontaine2015}).
  % Further, plants' roles [[how defined?]] within networks tend to be less phylogenetically constrained than those of 
  % animals~\citep{Rezende2007a,Chamberlain2014,Rohr2014,Vamosi2014,Lind2015}. [[GVDR is skeptical of this - check partiteness, size, type of role and add more detail to support.]] % I doubt this is really important, probably more trouble than it's worth.

@@

  We then estimated phylogenetic distances between species. To accomplish 
  this, we constructed a phylogenetic tree for our dataset based on a dated
  `mega-tree' of angiosperms~\citep{Zanne2014}. Some species were not included
  in the angiosperm mega-tree (largely ferns and tree ferns). For angiosperms,
  a sister taxon was identified using~\citet{APW} and the species added manually.
  Ferns, tree ferns, and a single club moss were added to the base of the tree.
  This means that closely-related non-angiosperm species appear to have very long 
  phylogenetic distances between them. For this reason, we excluded comparisons 
  between pairs of non-angiosperms from our analyses. As only two networks (both 
  herbivory networks) included more than one such species and non-angiosperms
  were always a small minority of any network, we do not 
  believe that omitting these comparisons has greatly affected our results.
  To obtain trees for each network, we 
  pruned the dated mega-tree to include only species in that network.

@@

  % [[It appears we switched to Jaccard and Sorenson dissimilarities - number of shared partners and number of shared partners*2/notshared.]]

\subsection*{Statistical analysis}

  We modelled the relationship between niche overlap and phylogenetic 
  distance using a logistic regression. We used both the numbers of shared 
  ($M_{ij}$) and non-shared ($U_{ij}$) partners as dependent variables and 
  centred, scaled phylogenetic distance as the independent variable. This 
  approach is conceptually similar to modelling successes and failures in a 
  binomial-distributed process. Accordingly, we assumed a binomial 
  distribution for residuals and used a logit link function to model the 
  probability $\omega_{ij}$ of plants $i$ and $j$ sharing an interaction 
  partner. Regressions of niche overlap and phylogenetic distance within 
  each network were fit using the R~\citep{R} base function ``glm'' and 
  took the form

    \begin{equation}
      \omega_{ij} \propto \beta_{distance}\delta_{ij} ,
      \label{basic}
    \end{equation}

  where $\delta_{ij}$ is the phylogenetic distance between plants $i$ and 
  $j$. The fixed effect of distance in this regression, $\beta_{distance}$, 
  can be understood as the change in log odds of sharing an interaction 
  partner per million-year change in phylogenetic distance.

  These separate regressions avoid the potential for confounding the effects
  of different relationships in different networks. As we also wished to 
  evaluate the overall trend across networks, we fit an additional
  regression of niche overlap and phylogenetic distance across all network 
  types. As well as the fixed effect of phylogenetic distance, this 
  regression included fixed effects of network type (pollination or 
  herbivory) and the interaction between phylogenetic network type and 
  random intercepts and slopes per network. This expanded regression was fit
  using the R~\citep{R} function ``glmer'' from package 
  \emph{lme4}~\citep{lme4} and took the form

    \begin{equation}
      \omega_{ij} \propto \beta_{distance} \delta_{ij} + \beta_{pollination} I_ij + \beta_{distance:pollination} \delta_{ij} I_ij ,
    \end{equation}

  where $I_{ij}=1$ when plants $i$ and $j$ are drawn from a 
  pollination network and $I_{ij}=0$ otherwise and all other symbols 
  are as above. Note that we only compared pairs of plants taken from the 
  same network. The fixed effects $\beta_{pollination}$ and $\beta_{distance:pollination}$ 
  are the change in intercept and slope of the log odds of sharing an 
  interaction partner, respectively relative to the baseline of 
  herbivory networks.

@@

  To test whether the relationship between phylogenetic distance and niche overlap depended on network size, we fit a general linear model for the slope of this relationship in the glm models against the number of plant pairs for which distances could be calculated (hereafter ``network size"), network type (again using herbivory networks as a baseline), and their interaction.

    \begin{equation}
      \beta_{distance} \propto \beta_{size} \eta_{N} + \beta_{pollination} I_N + \beta_{size:pollination} \eta_{N} I_N ,
    \end{equation}

  where $\eta_{N}$ is the number of plant pairs in network $N$ for which distances could be calculated, $I_N$ is an indicator equal to 1 if network $N$ is a pollination network and 0 otherewise.

  As the interaction between network type and network size was strong and opposite to the direction of the main effect of network size, we fit an additional general linear model using only data from pollination networks and including only the effect of network size (herbivory networks were the baseline in the full glm). Both models were fit using the R~\citep{R} base function ``glm".

@@

  For each plant family, within each network type, we then fit one of two 
  similar sets of models. If family $f$ was found in several networks of 
  the same type (e.g., several pollination networks), we fit a 
  mixed-effects logistic regression relating niche overlap to a fixed 
  effect of phylogenetic distance and a random effect for each network 
  using the R~\citep{R} function ``glmer'' from package
   \emph{lme4}~\citep{lme4}. If family $f$ was found in only one network
  (and therefore only one network type), we omitted the 
  network-level random effect and fit a logistic regression using the 
  R~\citep{R} base function ``glm''. These equations took the same form as equation~\ref{basic}.

@@

  \subsection*{Within-network conservation of niche overlap}

    Across all networks, more distantly-related plants were less likely to 
    share interaction partners ($\beta_{distance}$=-6.82, 
    $p$\textless0.001). Plants in pollination networks tended to share 
    fewer interaction partners overall, and the decrease in overlap with 
    increasing phylogenetic distance was steeper 
    ($\beta_{pollination}$=-1.44, $p$\textless0.001 and 
    $\beta_{distance:pollination}$=-18.5, $p$\textless0.001, respectively). 
    That is, a pair of plants in the same genus was more likely to share 
    interaction partners than a pair of plants in the same family in both 
    types of networks, but a pair of congeners would be less likely to 
    share pollinators than to share herbivores. [[Add predicted overlap values at extremes of distance?]] These trends may be partly 
    due to the greater proportion of specialist pollinators than specialist 
    herbivores. In our dataset, an average of 48\% (+/- 14) of pollinators 
    in a given web were extreme specialists (i.e., visited only one plant 
    species) compared to 29\% (+/- 29) of herbivores ($z$=5.62, df=68, 
    $P$\textless0.001 for a binomial regression of specialists and 
    generalists over network type).

    Despite these general trends, there was substantial variation between 
    pollination networks, with overlap of interaction partners decreasing 
    with increasing phylogenetic distance in some networks and increasing 
    in others (Fig.~\ref{within_network_regression}). Overlap of interaction partners decreased significantly with increasing phylogenetic distance in 7/11 herbivory networks and 33/59 pollination networks. In the remaining four herbivory networks and 25 of the 26 remaining pollination networks, overlap of interaction partners was not related to phylogenetic distance. Overlap of interaction partners increased with increasing phylogenetic distance in only a single pollination network.% (M_PL_026).
    The slope of the relationship between phylogenetic distance and overlap of interaction partners was related to the number of plant pairs in herbivory, but not pollination, networks. Larger herbivory networks had higher values of $\beta_{distance}$ ($\beta_{size}$=2.58$\times$10$^{-4}$, $p$=0.011 for the full glm; herbivory networks are the baseline). Plant-pollinator networks had higher (less negative) slopes overall ($\beta_{pollinator}$=0.306, $p$\textless0.001 compared to the intercept value of -0.434 for herbivory networks). Plant-pollinator networks moreover showed a much weaker relationship between network size and the strength of the overlap-distance relationship ($\beta_{pollination:size}$=-2.64$\times$10$^{-4}$, $p$=0.009). After refitting the glm to the pollination networks alone, there was no significant relationship between network size and the slope of the overlap-distance relationship ($\beta_{size}$=-5.91$\times$10$^{-6}$, $p$=0.572).

    [[to SI?]]
    Comparing the results in the observed networks to those obtained after permuting phylogenetic distances across pairs of plants, the observed slope of the relationship between phylogenetic distance and interaction partner overlap was always more extreme (i.e., always lesser or always greater) than that obtained in the permuted networks (Fig.~\ref{obs_vs_random}). Observed networks with a negative relationship between phylogenetic distance and overlap always had a more negative slope than that obtained from the permuted networks, while the 10 networks with positive relationships between phylogenetic distance and overlap always had more positive relationships than the permuted networks. This suggests that even in the networks with non-significant relationships between niche overlap and phylogenetic distance, the association between niche overlap and phylogenetic distance was not random.

    [[To SI.]]
    Comparing the permuted networks to permutations of the permuted networks, the slope obtained from the initial permuted network showed no clear relationship to the slopes obtained from 500 permutations of the permuted network. Averaged over the 1000 permutations of each observed network, the slope of the permuted network was more extreme than 48.1-51.3\% of the permutations of the permuted network. This confirms that shuffling phylogenetic distances between plant pairs destroys the relationship between distance and interaction partner overlap, and that further shuffling distances does not have a predictable effect.

@@

  % Regression of all families at once with RE's for network
    Taking all families together, the probability of species in the same family sharing interaction partners was not significantly related to phylogenetic distance ($\beta_{distance}$=-6.48, $p$=0.087). Pollination networks did not show a significantly different trend from the herbivory networks ($\beta_{distance:pollination}$=1.73, $p$=0.681). More closely-related pollinators did, however, tend to share fewer interaction partners ($\beta_{pollination}$=-0.776, $p$=0.007), similar to our within-network results above.

    Considering each family separately, the relationship between within-family niche overlap and phylogenetic distance varied widely in both pollination and herbivory networks. 
    For the families
    that were well represented in pollination networks, overlap decreased
    with increasing phylogenetic distance in 18 (Table~\ref{family_slopes_pp}).
    There was no significant relationship between overlap and phylogenetic distance in
    a further 15 plant families. Finally, the overlap between pairs of \emph{Polygonaceae}
    increased with increasing phylogenetic distance. 
    Of the seven plant families that were sufficiently well represented in herbivory 
    networks, overlap decreased with increasing phylogenetic distance in four 
    (Table~\ref{family_slopes_ph}; Fig.~\ref{within_family_regression}). Two
    families did not show significant relationships between phylogenetic distance and overlap,
    and in one family, \emph{Fabaceae}, overlap of interaction partners increased with 
    increasing phylogenetic distance.

    Models for two families did not converge. In both the \emph{Lauraceae}, (represented by four species in one pollination network) and the \emph{Sapindaceae} in herbivory networks (represented by five species in one network), one pair of species shared a single interaction partner while all other pairs did not share any interaction partners. The model for \emph{Sapindaceae} in pollination networks did converge.

@@

  ARC, DBS, GVDR, and NJB designed the research. ARC, MO, IN, and IMWH, JAT, and CJW collected published data. 
  ARC and GVDR performed the analyses. All authors contributed to the manuscript.

@@

  \begin{figure}[!h]
    \begin{center}
      \centerline{\includegraphics*[height=.75\textwidth]{Figures/dataplots/observed_vs_random.eps}}
    \end{center}
     \caption{The slopes of the mixed-effect logistic regression of pairwise niche overlap against phylogenetic distance was signficantly different from 0 for each network. Here we show, for herbivory (PH) and pollination (PP) networks, the observed slopes (depicted by green squares or purple diamonds, respectively). The maximum and minimum slopes obtained from 999 permutations of each network are depicted by thin, black lines. For both network types, the slopes obtained from permuted networks were always very close to 0.}
    \label{obs_vs_random} 
  \end{figure}
